\documentclass[addpoints, 10pt]{exam}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{systeme}
\usepackage{svg}
\usepackage{pgfplots}
\usepackage{bm}
\pgfplotsset{width=10cm,compat=1.9}
\title{MATH4.1EL Quiz 2 \\ Quadratic Equations in One Unknown \\ Time Limit: (40 minutes)}
\author{T Yeung}
\date{}

\renewcommand{\solutiontitle}{\noindent\textbf{\underline{Solution:}}\par\noindent}

\pagestyle{headandfoot}

\runningheadrule
\runningheader{Math 5.1EL}{Quiz 2}{14 Oct 2023}
\footer{}{Page \thepage\ of \numpages}{}
\begin{document}

\maketitle

\begin{center}
	\fbox{\fbox{\parbox{5.5in}{\centering
				Answer the questions in the spaces provided on the
				question sheets. If you do not know how to answer 
				a certain question, write down where you get stuck.
				Answers can be corrected to 3 significant figures
				if necessary.
	}}}
\end{center}
\vspace{0.1in}
\makebox[\textwidth]{Name, class, class no.:\enspace\hrulefill}
\vspace{0.3in}
\makebox[\textwidth]{Tutor’s name:\enspace\hrulefill}
\begin{questions}
	\question[3] 
	\begin{parts}
		\part[1] Solve the quadratic equation $x^2+3x+2=0$
		\part[2] Hence, solve the quadratic equation $2x^2+3x+1=0$
	\end{parts}
	\question[12]
	\begin{parts} \part Let $\alpha$, $\beta$ be the roots of $x^2+px+q=0 \quad (*)$, where $p$ and $q$ are constants. Find, in terms of $p$ and $q$.
		\begin{subparts}
			\subpart[2] $\alpha^2+\beta^2$
			\subpart[2] $\alpha^3+\beta^3$
			\subpart[2] $(\alpha^2-\beta-1)(\beta^2-\alpha-1)$.
		\end{subparts}
		\part[2] If the square of one root of (*) minus the other root equals $1$, use (a), or otherwise, to show that $q^2-3(p-1)q+(p-1)^2(p+1)=0 \quad (**)$
		\part[2] Find the range of values of $p$ such that the quadratic equation $(**)$ has real roots.
		\part[2] Suppose $k$ is a real constant. If the square of one root of $4x^2+5x+k=0$ minus the other root equals $1$, use the result in $(b)$, or otherwise, to find the value of $k$.
	\end{parts}
	\question[4] The quadratic equation $9x^2-(k+1)x+1=0 \quad (*)$ has equal roots.
	\begin{parts}
		\part Find the two possible values of the constant $k$.
		\part If $k$ takes the negative value obtained, solve equation $(*)$.
	\end{parts}
	\question[4] $\alpha$ and $\beta$ are the roots of the quadratic equation $kx^2-4x+2k=0$, where $k (k \neq 0)$ is a constant. Express the following in terms of $k$.
	\question[4] The equation $x^2+4x+p=0$, where $p$ is a real constant, has distinct real roots $\alpha$ and $\beta$.
	\begin{parts}
		\part Find the range of values of $p$
		\part If $\alpha^2+\beta^2+\alpha^2\beta^2+3(\alpha+\beta)-19=0$, find the values of $p$.
	\end{parts}
\end{questions}
\end{document}
